Demystifying the Computer Science Ph.D. Admission in the US

### 2023-06-09 phd-cs-us.tex
updates csrankings table

\begin{abstract}
Having been involved in PhD admissions for many years and after
numerous interaction with  students, I've
realized that \emph{international} students, especially Vietnamese, lack a clear understanding of
the Computer Science PhD admission process at US universities. This confusion not only
discourages students from applying but also creates the perception that
getting admitted is difficult compared to CS PhD programs in other countries.

@@

First, I want to emphasize that PhD students in Computer
Science \emph{do not} need to worry about funding, especially at R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} to support your study, including tuition,
health insurance, and 9-month stipend. Moreover, depending on the university,
you may even receive additional benefits like summer salary. Note that
funding is typically more available for PhD students than 
Masters and especially undergraduate studies, which typically have no funding at all. \S\ref{sec:funding} provides more details on funding.

@@

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: One of the reasons I create this post is that several of my colleagues are interested in 
recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (compared to other countries). In general the number of
PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
\end{tcolorbox}

@@

In many cases, the admission committee will involve junior faculty (e.g., assistant professors) in the department. While this requires  a significant time investment, it provides junior faculty the opportunities to recruit students. The chair of the committee will be a senior professor, but they will not review individual applications and instead assign them to committee members. The chair will look at various factors such as research interests or mentioning faculty names, and assign the applications to appropriate faculty.

@@

\subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
  Below is the 2023 Rankings from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on top CS conferences, as of April 2023.

\begin{itemize}
\item Carnegie Mellon University 
\item Univ. of Illinois at Urbana-Champaign 
\item Univ. of California - San Diego  
\item Massachusetts Institute of Technology 
\item Stanford University         
\item University of Michigan  
\item University of Washington 
\item Cornell University  
\item Univ. of California - Berkeley  
\item Georgia Institute of Technology 
\item Northeastern University  
\item University of Maryland - College Park 
\item University of Wisconsin - Madison  
\item Purdue University  closed chart
\item University of Texas at Austin  
\item Columbia University  
\item University of Pennsylvania 
\item New York University  
\item Princeton University 
\item University of Massachusetts Amherst  
\item Univ. of California - Los Angeles
\item University of Chicago 
\item Stony Brook University
\item Univ. of California - Santa Barbara 
\item Rutgers University  
\item Univ. of California - Irvine 
\item University of Southern California 
\item Duke University 
\item Univ. of California - Riverside
\item Pennsylvania State University 
\item Northwestern University 
\item Ohio State University 
\item George Mason University 
\item Harvard University  
\item Brown University  
\item University of Utah 
\item Texas A\&M University
\item Univ. of California - Santa Cruz  
\item Yale University  
\item Boston University 
\item University at Buffalo
\item University of Colorado Boulder 
\item North Carolina State University
\item Rice University 
\item University of Illinois at Chicago  
\item University of North Carolina 
\item University of Virginia 
\item Arizona State University 
\item University of Minnesota 
\item Virginia Tech  
\item  Oregon State University  
\item Univ. of California - Davis
\end{itemize}

>>>>

\begin{abstract}
Having been involved in PhD admissions for many years and having
numerous interaction with  students, I've
realized that \emph{international} students, especially \textbf{Vietnamese}, lack a clear understanding of
the Computer Science PhD admission process at US universities. This confusion not only
discourages students from applying but also creates the perception that
getting admitted is difficult compared to CS PhD programs in other countries.

@@

First, I want to emphasize that PhD students in Computer
Science \emph{do not} need to worry about funding, especially at good R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} to support your study, including tuition,
health insurance, and stipend. Moreover, depending on the university,
you may even receive additional benefits such as summer salary and laptops. Note that
funding is typically more available for PhD students than 
Masters and especially undergraduate studies, which typically have no funding. \S\ref{sec:funding} provides more details on funding.

@@

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: One of the reasons I create this post is that several of my colleagues are interested in 
recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 3 applications from Vietnam each year). In general the number of
PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
\end{tcolorbox}

@@

In many cases, the admission committee can involve current students and assistant professors in the department. This provides junior faculty the opportunities to recruit students. The chair of the committee will be a senior professor, but they likely will not review individual applications and instead assign them to committee members. The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.

@@

\subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
  Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on CS conferences.

  \begin{table}
    \centering
    \small
    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, April 2023)}\label{tab:ranking}
  \begin{tabular}{rl|rl}
    \toprule
    1 & Carnegie Mellon University & 26 & Univ. of California - Irvine \\
    2 & Univ. of Illinois at Urbana-Champaign  & 27 & University of Southern California \\
    3 & Univ. of California - San Diego & 28 &  Duke University \\
    4 & Massachusetts Institute of Technology & 29 & Univ. of California - Riverside\\
    5 & Stanford University         & 30 & Pennsylvania State University \\
    6 & University of Michigan - Ann Arbor & 31 & Northwestern University \\
    7 & University of Washington   & 32& Ohio State University \\  
    8 & Cornell University       &33 & George Mason University \\
    9 & Univ. of California - Berkeley  &34& Harvard University  \\
    10 & Georgia Institute of Technology &35& Brown University  \\
    11 & Northeastern University  &36& University of Utah \\
    12 & University of Maryland - College Park &37& Texas A\&M University\\
    13 & University of Wisconsin - Madison  &38& Univ. of California - Santa Cruz  \\
    14 &   Purdue University  &39 & Yale University  \\
    15 & University of Texas at Austin  &40& Boston University \\
    16 &   Columbia University  &41 & University at Buffalo\\
    17 & University of Pennsylvania &42& University of Colorado-Boulder \\
    18 & New York University  & 43& North Carolina State University\\
    19 & Princeton University & 44& Rice University \\
    20 & University of Massachusetts Amherst &45& University of Illinois at Chicago  \\
    21 & Univ. of California - Los Angeles &46& University of North Carolina \\
    22 & University of Chicago &47& University of Virginia \\
    23 & Stony Brook University &48& Arizona State University \\
    24 & Univ. of California - Santa Barbara &49& University of Minnesota \\
    25 & Rutgers University &50& Virginia Tech  \\
    \bottomrule
    \end{tabular}
\end{table}
